---
title: "Difference-in-Differences"
subtitle: "The very basics"
date: 2026-07-10
author: "G. Brueckmann"
format: 
  revealjs:
    code-summary: "Display R code"
    code-tools: true
    embed-resources: true
---
  
## What is  *Difference-in-Differences*?
  
::: {.callout-important collapse="true"}
### The basic idea of difference-in-differences
Using the **interaction** of *time* and (quasi-)random *assignment to the intervention* to estimate the causal effect of the intervention.
:::
  
  <!-- ## The Set-Up in R -->
  
```{r}
#| label: setup 
#| code-summary: "Display R code loading relevant R packages"
#| echo: true
#| code-fold: true
#| message: false
#| warning: false

library(tidyverse) # to do data wrangling
library(tinytable) # to report tables nicely

library(causaldata) # to get data accompanying Huntington-Klein's `The Effect` book¨
library(modelsummary) # to report results

library(fixest) # for fixed-effects estimation 
```

## Example 1: Cholera in London

In [The Effect (Chapter 18)](https://theeffectbook.net/ch-DifferenceinDifference.html) Huntington-Klein cites *John Snows' (1855)* study on cholera transmission (through faecal matters) in London's drinking water.
Between 1849 and 1854 Lambeth moved their well upstream, thus, the water becomes cleaner.

```{r}
#| label: tbl-lndtable
#| tbl-cap: Death rates by areas with different water suppliers
#| echo: true
#| code-fold: true
#| message: false
#| warning: false



lnddata <- tibble::tribble(~'Water Supplier', ~ '1849',  ~ '1854',
                           'Non-Lambeth', 134.9, 146.6,
                           'Lambeth', 130.1, 84.9)


tt(lnddata, width = c(0.45, 0.1, 0.1)) |>
  format_tt(replace = "") |>
  style_tt(j = c(2:3), align = "c") |>
  group_tt(j = list("Death Rates" = c(2, 3)))  |>
  theme_tt("revealjs")

```

------------------------------------------------------------------------
  
### Ex1: Differences between water suppliers {auto-animate="true"}
  
```{r}
#| label: tbl-lnddiffbetweenwells
#| tbl-cap: Differences in death rates between water suppliers (at each point in time) 
#| echo: true
#| code-fold: true
#| message: false
#| warning: false


diff_row <- lnddata |>
  summarise(
    `Water Supplier` = "Difference (between suppliers)",
    `1849` = `1849`[1] - `1849`[2],
    `1854` = `1854`[1] - `1854`[2]
  )

diffs_sup <- bind_rows(lnddata, diff_row)

tt(diffs_sup,width = c(0.45, 0.1, 0.1)) |>
  format_tt(replace = "") |>
  style_tt(j = c(2:3), align = "c") |>
  style_tt(i = 3, bold = TRUE) |> 
  group_tt(j = list("Death Rates" = c(2, 3)))  |>
  theme_tt("revealjs")
```

------------------------------------------------------------------------
  
### Ex1: Differences over time {auto-animate="true"}
  
```{r}
#| label: tbl-lnddiffovertime
#| tbl-cap: Differences in death rates over time (each supplier)
#| echo: true
#| code-fold: true
#| message: false
#| warning: false

diffs_time <- diffs_sup  |>
  mutate(
    `Difference\n(1854-1849)` = 
      if_else( # use if else to only calculate differences over time per supplier
        `Water Supplier` %in% c("Non-Lambeth", "Lambeth"),
        `1854` - `1849` ,
        NA
      )
  )

tt(diffs_time, width = c(0.45, 0.1, 0.1, 0.2))|>
  group_tt(j = list("Death Rates" = c(2, 3)))  |>
  format_tt(replace = "") |>
  style_tt(j = c(2:3), align = "c") |>
  style_tt(j = 4, bold = TRUE) |> 
  theme_tt("revealjs")

```

------------------------------------------------------------------------
  
### Ex1: Compute the difference-in-differences {auto-animate="true"}
  
  
```{r}
#| label: tbl-lnddiffindiff
#| tbl-cap: Differences in differnces of death rates over time and between suppliers
#| echo: true
#| code-fold: true
#| message: false
#| warning: false


lnd_diffindiffs <- diffs_time  |>
  mutate(
    `Difference\n(1854-1849)` = 
      `1854` - `1849`
  )

tt(lnd_diffindiffs, width = c(0.45, 0.1, 0.1, 0.2))|>
  group_tt(j = list("Death Rates" = c(2, 3)))  |>
  format_tt(replace = "") |>
  style_tt(j = c(2:3), align = "c") |>
  style_tt(i = 3, j = 4, bold = TRUE) |> 
  style_tt(i = 3, j = 1, bold = TRUE) |> 
  theme_tt("revealjs")



lnd_did_pe <- lnd_diffindiffs  |> filter(`Water Supplier` == "Difference (between suppliers)")  |> select(`Difference\n(1854-1849)`)  |> pull()
```


<!-- This text appears later. -->
::: {.fragment}
$\Rightarrow$ Differences in death rates between suppliers (1854-1849) (or differences in death rates over time between suppliers) gave you the *diff-in-diffs* point estimate of **`{r} lnd_did_pe`**.
:::
  
  
------------------------------------------------------------------------
  
### Ex1: Graphical Representation {.nostretch}
  
```{r}
#| label: createfigure
#| code-summary: "Display R code to create the figure"
#| message: false
#| warning: false
#| code-fold: true


# make data from wide to long format for plotting
lnddata_long <- lnddata |>
  pivot_longer(
    cols = -`Water Supplier`,
    names_to = "Year",
    values_to = "Death_Rate",
    names_transform = list(Year = parse_number)### only keep the year not any text 
  )  |> rename(Supplier = `Water Supplier`)

londonplot <- 
  ggplot(lnddata_long, aes(Year, Death_Rate)) +
  geom_point(aes(
    colour = Supplier,
    # outline colour
    shape  = Supplier
  ), size = 9) +
  geom_smooth(
    aes(colour = Supplier),
    # one line per supplier
    method  = "lm",
    formula = y ~ x,
    linewidth = 0.6,
    # smaller value = thinner line
    alpha   = 0.1 ,
    # set line transparency
    show.legend = FALSE
  ) +
  scale_shape_manual(
    values = c(
      "Lambeth" = "square cross",
      # treated gets a square
      "Non-Lambeth" = "circle plus"     # circle for control
    ),
    labels = c( "Non-Lambeth"     = "Non-Lambeth", "Lambeth" = "Lambeth")
  )   +
  # set outline colours manually and rename
  scale_colour_manual(
    values = c(
      "Lambeth" = "#5DC863",
      # viridis light green-ish
      "Non-Lambeth"     = "#440154"   # viridis purple-ish
    ),
    labels = c( "Non-Lambeth"     = "Non-Lambeth", "Lambeth" = "Lambeth")
  ) +
  scale_x_continuous(
    labels = c("1849", "1854"),
    breaks = c(1849, 1854),
    minor_breaks = seq(1849, 1854, 1)
  ) +
  labs(y = "Death rates")   +
  coord_fixed(ratio = 0.08,
              expand = TRUE,
              reverse = "y")   +### make the plot more squared
  theme_minimal(base_size = 20) +
  theme(
    panel.grid.minor = element_blank(),
    # remove grey fill
    strip.background = element_blank(),
    panel.border = element_blank(),
    plot.title = element_blank(),
    axis.line.y.left = element_line(colour = "black"),
    axis.line.x.bottom  =   element_line(colour = "black"),
    axis.text = element_text(size = 18, colour = "black"),
    axis.title.y = element_text(size = 18, colour = "black"),
    #define the legend
    legend.text = element_text(size = 18, colour = "black"),
    legend.title = element_blank(),
    legend.key = element_rect(fill = NA, color = NA),
    #define the legend inside the plot area
    legend.position = "inside",
    legend.justification = c("left", "top")
  )


londonplot_sidelegend <- ggplot(lnddata_long, aes(Year, Death_Rate)) +
  geom_point(aes(
    colour = Supplier,
    # outline colour
    shape  = Supplier
  ), size = 9) +
  geom_smooth(
    aes(colour = Supplier),
    # one line per supplier
    method  = "lm",
    formula = y ~ x,
    linewidth = 0.6,
    # smaller value = thinner line
    alpha   = 0.1 ,
    # set line transparency
    show.legend = FALSE
  ) +
  scale_shape_manual(
    values = c(
      "Lambeth" = "square cross",
      # treated gets a square
      "Non-Lambeth" = "circle plus"     # circle for control
    ),
    labels = c( "Non-Lambeth"     = "Non-Lambeth (Control)", "Lambeth" = "Lambeth (treated in 1854)")
  )   +
  # set outline colours manually and rename
  scale_colour_manual(
    values = c(
      "Lambeth" = "#5DC863",
      # viridis light green-ish
      "Non-Lambeth"     = "#440154"   # viridis purple-ish
    ),
    labels = c( "Non-Lambeth"     = "Non-Lambeth (Control)", "Lambeth" = "Lambeth (treated in 1854)")
  ) +
  scale_x_continuous(
    labels = c("1849\n(pre-treatment)", "1854\n(post-treatment)"),
    breaks = c(1849, 1854),
    minor_breaks = seq(1849, 1854, 1)
  ) +
  labs(y = "Death rates")   +
  coord_fixed(ratio = 0.08,
              expand = TRUE,
              reverse = "y")   +### make the plot more squared
  theme_minimal(base_size = 20) +
  theme(
    panel.grid.minor = element_blank(),
    # remove grey fill
    strip.background = element_blank(),
    panel.border = element_blank(),
    plot.title = element_blank(),
    axis.line.y.left = element_line(colour = "black"),
    axis.line.x.bottom  =   element_line(colour = "black"),
    axis.text = element_text(size = 18, colour = "black"),
    axis.title.y = element_text(size = 18, colour = "black"),
    #define the legend
    legend.text = element_text(size = 18, colour = "black"),
    legend.title = element_blank(),
    legend.key = element_rect(fill = NA, color = NA),
    legend.position = "right")  

```
::: {data-id="myfig"}
```{r}
#| label: fig-lnd
#| message: false
#| warning: false
#| code-fold: true
#| fig-align: center
#| fig-cap: The Cholera in London example, where Lambeth is considered *treated* in 1854.
#| fig-height: 5
#| fig-width: 5
#| out-width: 55%
londonplot
```
::: 
  
------------------------------------------------------------------------
  
### Ex1: Suggestions for additional exercises {auto-animate="true" .nostretch}
::: {data-id="myfig"}
```{r}
#| label: lnd-smaller
#| message: false
#| warning: false
#| code-fold: true
#| fig-align: center
#| fig-height: 5
#| fig-width: 5
#| out-width: 30%
londonplot
```
:::
::: {.callout-tip collapse="true"}
#### Suggestion: As additional exercises ...
- visualise all differences computed in @tbl-lnddiffindiff in the @fig-lnd
- write the corresponding regression equation using time and treatment dummies 
- visualise the resulting different $\beta$ coefficients in @fig-lnd
:::
  
  
## Example 2: Organ donation system 
Now, consider California's July 2011 switch from an opt-in to an active choice
organ donation system. 

This example also stems from [The Effect (Chapter 18)](https://theeffectbook.net/ch-DifferenceinDifference.html) by Huntington-Klein.

------------------------------------------------------------------------
  
### Ex2: The effect of California's policy switch
::: {.callout-tip collapse="true"}
#### Suggestion:
Look at @tbl-californiaex on the next slide and recall why there is only one coefficient in the regression. You might want to look at the R Code.
:::


```{r}
#| label: tbl-californiaex
#| tbl-cap: Two-way fixed effects using the `feols` package
#| echo: true
#| code-fold: true
#| message: false
#| warning: false


# import data
od <- causaldata::organ_donations

# Treatment variable
od <- od |>
  mutate(Treated = State == 'California' &
           Quarter %in% c('Q32011', 'Q42011', 'Q12012'))

# feols clusters by the first
# fixed effect by default, no adjustment necessary
clfe <- feols(Rate ~ Treated | State + Quarter, data = od)
modelsummary(
  clfe,
  stars = TRUE,
  vcov = ~ State,  # to display the same SE as in Table 18.2 of The Effect
  gof_omit = 'AIC|BIC' ,
  escape = FALSE
) |> theme_tt("revealjs")

```

## Important R functions & packages: {auto-animate="true"}
`feols` from `fixest` and  `felm` from `lfe`.

::: {.callout-warning collapse="true"}
#### Consider the different `lfe` syntax:
```r
library(lfe) 

twfe <- list(
  "Raw SE" <- felm(Rate ~ Treated | State + Quarter | 0 , od) ,  
  # Note: there are only two vertical lines
  # (around the fixed-effects)
  "Clustered SE" <- felm(Rate ~ Treated | State + Quarter | 0 | State, od) 
  # NORE: the third vertical line 
  # for clustering the SEs (by State) at the very end!
)
# Two see the differences, consider using modelsummary for display
modelsummary(twfe,
             stars = TRUE,
             gof_omit = 'AIC|BIC' ,
             escape = FALSE) |> 
  theme_tt("revealjs")
```
:::


## Important R functions & packages: {auto-animate="true"}



::: callout-note
### Note that `feols` from `fixest` works better with `modelsummary` (called after estimation for display) than `lfe` (which I want to warn against using). 
Note that using `felm` from the `lfe` package gives you the same point estimate as @tbl-californiaex but `modelsummary` (called after estimation for display) does not let you cluster the SEs the same way as in @tbl-californiaex (as `modelsummary` relies on the `sandwich` package, which can't handle `lfe` variance-covariance-matrix estimates, see <https://github.com/jepusto/clubSandwich/issues/40>). This means, use the second specification using `| SE_CLUSTER_VARIABLE` if you use `felm` from the `lfe` package.
:::  
## Additional Resources
  
### Basics:
  
  <https://statisticsglobe.com/fixed-effects-linear-regression>
  
  <https://rpubs.com/corinne-riddell/guide-to-did-estimators>
  
### Advanced:
  
  <https://asjadnaqvi.github.io/DiD/>
  
  <https://psantanna.com/did-resources/>
  
  (has many resources!)
